Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2025-01-31 article.tex
Add a script to automatically update the speedup plots

The current work focusses on vector problems, where the governing momentum equation is formulated in terms of the unknown displacement solution vector.
In this case, Equation \ref{eq:J} refers to the individual scalar components of the residual, solution, and Jacobian.
That is, for 3-D analyses, the residual takes the form
\begin{eqnarray}
	\bb{R}(\bb{u}) = \left\{ R_1^x, R_1^y, R_1^z, R_2^x, R_2^y, R_2^z, ..., R_n^z \right\}
\end{eqnarray}
and the solution takes the form
\begin{eqnarray}
	\bb{u} = \left\{ u_1^x, u_1^y, u_1^z, u_2^x, u_2^y, u_2^z, ..., u_n^z \right\}
\end{eqnarray}
In practice, it is often more practical and efficient to form and store the residual, solution and Jacobian in a \emph{blocked} manner, where the residual and solution can be considered as vectors of vectors.
Similarly, the Jacobian can be formed in terms of sub-matrix block coefficients.

@@

\hl{magSf in OF/ d3 weights}

@@

To achieve this, the cell-centred displacement gradients are determined using an inverse-distance-weighted first-neighbours least squares method \citep{Jasak1996},

@@

		&=&\sum_{f_i \in \mathcal{F}^{\text{int}}_P} \frac{\bb{G}^{-1}_P \cdot \bb{d}_{f_i}}{\bb{d}_{f_i} \cdot \bb{d}_{f_i}}  \otimes \left(\bb{u}_{N_{f_i}} - \bb{u}_P \right) \notag \\
		&&+ \sum_{d_i \in \mathcal{F}^{\text{disp}}_P} \frac{ \bb{G}^{-1}_P \cdot \bb{d}_{d_i} }{\bb{d}_{d_i} \cdot \bb{d}_{d_i}} \otimes \left(\bb{u}_{d_i} - \bb{u}_P \right) \notag \\
		&&+ \sum_{s_i \in \mathcal{F}^{\text{symm}}_P} \frac{\bb{G}^{-1}_P \cdot \bb{d}_{s_i}}{\bb{d}_{s_i} \cdot \bb{d}_{s_i}} \otimes \left(\bb{R}_{s_i} \cdot \bb{u}_P - \bb{u}_P \right)

@@

	 \sum_{{f_i} \in \mathcal{F}^{\text{int}}_P} \frac{\bb{d}_{f_i} \otimes \bb{d}_{f_i}}{\bb{d}_{f_i} \cdot \bb{d}_{f_i}}
	 +  \sum_{{d_i} \in \mathcal{F}^{\text{disp}}_P} \frac{\bb{d}_{d_i} \otimes \bb{d}_{d_i}}{\bb{d}_{d_i} \cdot \bb{d}_{d_i}}
	 +  \sum_{{s_i} \in \mathcal{F}^{\text{symm}}_P} \frac{\bb{d}_{s_i} \otimes \bb{d}_{s_i}}{\bb{d}_{s_i} \cdot \bb{d}_{s_i}}

@@

where the $\bb{G}^{-1}_P \cdot \bb{d}/(\bb{d}\cdot \bb{d})$ vectors are purely a function of the mesh and can be computed once and stored.

@@

first-order accurate gradients are sufficient to preserve second-order accuracy of the cell-centre displacements. \hl{OF uses magSf}

@@

So, this Rhie-Chow stabilisation can be seen to take the form of a \emph{jump} term; that is, it is a function of the jump in solution between neighbouring cells.
The two states, $\bb{u}_P^*$ and $\bb{u}_{N_{f_i}}^*$, are evaluated halfway between $P$ and $N_{f_i}$, and it may not be at a face centre; however, as noted by \citet{Nishikawa2010}, this will not affect the second-order accuracy of the method; consequently, the user is free to choose the value of $\alpha$ for stabilisation and accuracy purposes.
% but it doesn't matter because this team is only stabilisation and it vanishes in the grid refinement.
%As long as it vanishes for linear functions, second-order accuracy is preserved. It means that the parameter alpha can be chosen solely for the purpose of stabilization.

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Segregated Solution Algorithm} 
\label{sec:seg_alg}
%%--------------------------------------------------------------------------------------------------------------------%%
\hl{include alpha and check its effect}
The classic segregated solution algorithm can be viewed as a quasi-Newton method, where an approximate Jacobian is derived from the inertia term and a compact-stencil discretisation of a diffusion term:
%The surface forces Laplacian term (first term on the right-hand side of Equation \ref{eqn:MomentumImplicitExplicit}) is discretised using central differencing with over-relaxed non-orthogonal correction \cite{demirdzic_finite_1993, jasak_application_2000, cardiff_development_2014, cardiff_large_2014, cardiff_lagrangian_2017}:
\begin{eqnarray} \label{eq:diffusion}
	\tilde{\bb{J}} &=& \frac{\partial}{\partial \bb{u}} \left[ \oint_{\Gamma_P} \bar{K} \, \bb{n} \cdot \bb{\nabla} \bb{u} \; d\Gamma_P
	 \; -\;  \int_{\Omega_P} \rho \frac{\partial^2 \bb{u} }{\partial t^2} \, d\Omega_P \right]
\end{eqnarray}
The inertia term is discretised as described in Equations \ref{eq:inertia} and \ref{eq:inertia2}, while the diffusion term is discretised in the same manner as the compact-stencil component of the Rhie-Chow stabilisation term:
\begin{eqnarray}
	\oint_{\Gamma_P} \bar{K} \, \bb{n} \cdot \bb{\nabla} \bb{u} \; d\Gamma_P &\approx&
		\sum_{f_i \in \mathcal{F}^{\text{int}}_P}  \bar{K}
		\left|\bb{\Delta}_{f_i} \right| \frac{ \bb{u}_{N_{f_i}} - \bb{u}_P}{\left|\bb{d}_{f_i}\right|}    \left|\bb{\Gamma}_{f_i}\right| \notag \\
	&&+  \sum_{d_i \in \mathcal{F}^{\text{disp}}_P}  \bar{K}
%		\left|\bb{\Delta}_{d_i} \right| \left( \ \frac{ \bar{\bb{u}}_{d_i} - \bb{u}_P}{\left|\bb{d}_{d_i}\right|}  \right) 
		\left|\bb{\Delta}_{d_i} \right| \frac{ \bar{\bb{u}}_{d_i}  - \bb{u}_P}{\left|\bb{d}_{d_i}\right|} 
		\left|\bb{\Gamma}_{d_i}\right| \notag \\
	&&+ \sum_{s_i \in \mathcal{F}^{\text{symm}}_P}  \bar{K}
		\left|\bb{\Delta}_{s_i} \right| \frac{ \bb{R}_{s_i} \cdot \bb{u}_{P} - \bb{u}_P}{\left|\bb{d}_{s_i}\right|}
		\left|\bb{\Gamma}_{s_i}\right|
\end{eqnarray}
When a diffusion term is typically discretised using the cell-centre finite volume method, non-orthogonal corrections are included in a deferred correction manner to preserve the order of accuracy on distorted grids.
However, in the current quasi-Newton method form, the approximate Jacobian's exact value does not affect the final converged solution, but only the convergence behaviour.
Consequently, non-orthogonal corrections are not included in the approximate Jacobian here. However, grid distortion is appropriately accounted for in the calculation of the residual.
Nonetheless, as a result, it is expected that the convergence behaviour of the segregated approach may degrade as mesh non-orthogonality increases.
%Similarly, the prescribed displacements from the $\mathcal{F}^{\text{disp}}_P$ boundary faces does not affect the Jacobian.

The linearised system (Equation \ref{eq:Seg}) is formed for each cell in the domain, resulting in a system of algebraic equations:
\begin{eqnarray} \label{eq:SegSys}
    \left[ \bb{\tilde{J}} \right]  \; \left[ \delta \bb{u} \right] = - \left[\bb{R}(\bb{u}_k)\right]
\end{eqnarray}
where $\left[ \bb{\tilde{J}} \right]$ is a symmetric, $M \times M$ stiffness matrix, where $M = 3|\mathcal{P}|$ in 3-D and $M = 2|\mathcal{P}|$ in 2-D.
If $\Delta t < \infty$ or $\mathcal{F}^{\text{disp}}_P \neq \emptyset$, matrix $\left[ \bb{\tilde{J}} \right]$ is strongly diagonally dominant; otherwise, it is weakly diagonally dominant.
The block ($3\times3$ for 3-D, $2\times2$ for 2-D) diagonal coefficient for cell $P$ (row $P$, column $P$) can be expressed as
\begin{eqnarray}
	 \left[ \tilde{\bb{J}}\right]_{PP} &=&
		- \sum_{f_i \in \mathcal{F}^{\text{int}}_P}  \bar{K}
		\frac{ \left|\bb{\Delta}_{f_i} \right| }{\left|\bb{d}_{f_i}\right|}    \left|\bb{\Gamma}_{f_i}\right| \textbf{I} 
	    \quad-  \sum_{d_i \in \mathcal{F}^{\text{disp}}_P}  \bar{K}
%		\left|\bb{\Delta}_{d_i} \right| \left( \ \frac{ \bar{\bb{u}}_{d_i} - \bb{u}_P}{\left|\bb{d}_{d_i}\right|}  \right) 
		 \frac{ \left|\bb{\Delta}_{d_i} \right| }{\left|\bb{d}_{d_i}\right|} 
		\left|\bb{\Gamma}_{d_i}\right| \textbf{I} \notag \\
	 &&\quad - \sum_{s_i \in \mathcal{F}^{\text{symm}}_P}  \bar{K}
		 \frac{ \left|\bb{\Delta}_{s_i} \right|}{\left|\bb{d}_{s_i}\right|}
		\left|\bb{\Gamma}_{s_i}\right|  \bb{n}_{s_i} \otimes \bb{n}_{s_i} 
	\quad - \quad \frac{9}{4}  \frac{\rho_P \Omega_P}{\Delta t^2} \textbf{I}
\end{eqnarray}
while the off-diagonal coefficients (row $P$, column $Q$) can be expressed as
\begin{eqnarray}
	\left[\tilde{\bb{J}}\right] _{PQ} &=&
		\bar{K} \frac{ \left|\bb{\Delta}_{f_{PQ}} \right| }{\left|\bb{d}_{f_{PQ}}\right|}    \left|\bb{\Gamma}_{f_{PQ}}\right| \textbf{I} 
\end{eqnarray}
where subscript $PQ$ indicates a quantity associated with the internal face $f$ shared between cells $P$ and $Q$.
%where $N_f$ represents the set of faces $f$ in cell $P$, and neighbouring cell centre $N_f$ shares face $f$ with the cell $P$.
%The over-relaxed orthogonal vector $\bb{\Delta}_f = \frac{\bb{d}_f}{\bb{d}_f \cdot \bb{n}_f}$ 
%The non-orthogonal correction vector $\bb{k}_f=\bb{n}_f-\bb{\Delta}_f$.
% where $\bb{n}_f$ is the outward-facing unit normal to the face $f$.
%Vector $\bb{d}_f$ connects the centre of cell $P$ with the centre of cell $N_f$ in the updated configuration.
%The first term on the right-hand side is treated implicitly, while the second term - representing non-orthogonal corrections at the face - is treated in a deferred correction manner.

@@

In literature, the segregated solution algorithm is typically formulated in terms of the total displacement vector (or its difference between time steps) as the primary unknown; in contrast, in the quasi-Newton interpretation presented here, the primary unknown is the correction to the displacement vector, which goes to zero at convergence.
Nonetheless, both approaches are equivalent and neither formulation might be expected to display superior performance.

@@

%\hl{3}
%- we use approx J to form precon
%- many precon used in lit, we will consider ILU($k$) and MG, where MG is expected to be better, but also LU since direct solvers are popular in FE solid mechanics
In the current work, we propose to use the compact-stencil approximate Jacobian from the segregated algorithm $\tilde{\bb{J}}$ as the preconditioning matrix $\bb{P}$ for the preconditioned Jacobian-free Newton-Krylov method.
%\hl{Comment on literature: used before but not yet for solids}.
This preconditioning approach can be considered as a ``physics-based" preconditioner in the classifications of \citet{Knoll2004}.
The approach is conceptually similar to an approximation of the Jacobian of a higher-order large-stencil scheme by a compact-stencil lower-order scheme.
A benefit of the proposed approach is that existing segregated frameworks can re-use their existing matrix assembly and storage implementations.
Concretely, the Jacobian-free Newton-Krylov method requires only a procedure for forming this preconditioning matrix and a procedure for explicitly evaluating the residual.
Both routines are easily implemented -- and are likely already available -- in an existing segregated framework.
The only additional required procedure is an interface to an existing Jacobian-free Newton-Krylov implementation.
%"The motivation behind this approach is that there exist numerous, legacy algorithms to solve nonlinear systems, both IVPs and BVPs. These algorithms typically were developed with some insight into the time scales or physical behavior of the problem. As a benefit of this insight, a reduced implicit system, or a sequence of segregated explicit or implicit systems may be solved in place of the fully coupled system. "
In the current work, the PETSc package \cite{PETSc} is used as the nonlinear solver, driven by a finite volume solver implemented in the solids4foam toolbox \citep{Cardiff2018, Tukovic2018} for OpenFOAM toolbox \citep{Weller1998} (version OpenFOAM-v2312).

@@

%\hl{5 - oversolving}
%Comment on over-solving => also applies to the segregated system.
%This could be a parameter we look at.
A final comment on the Jacobian-free Newton-Krylov solution algorithm is the potential importance of \emph{oversolving}.
Here, oversolving refers to solving the linear system to too tight a tolerance during the early Newton iterations, essentially wasting time when the solution is far from the true solution.
In addition, some authors %\cite{See164_and_176_in_Knoll2004}
\cite{Knoll2004} have shown Newton convergence to be worse when the earlier iterations are solved to too tight a tolerance.
The concept of oversolving also applies to segregated solution procedures and has been well-known since the early work of Demird\v{z}i\'{c} and co-workers \cite{Demirdzic1995}, where the residuals are typically reduced by one order of magnitude in the inner linear system.
%The optimal choice of residual reduction for a Jacobian-free Newton-Krylov finite volume solid mechanics procedure is explored in Section \ref{sec:test_cases}. \hl{Check: do we examine this?}
In the current work, the residual is reduced by one order of magnitude in the segregated approach and three orders in the Jacobian-free Newton-Krylov approach.
\hl{Try for EW plastic: $-snes_ksp_ew -snes_ksp_ew_version 2 -snes_ksp_ew_rtol0 1e-4$}

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Checking Convergence}
%%--------------------------------------------------------------------------------------------------------------------%%
\hl{sTol vs fTol vs aTol}
\hl{sTol: not equivalent in JFNK and Picard}

>>>>

The current work focuses on vector problems, where the governing momentum equation is formulated in terms of the unknown displacement solution vector.
In this case, Equation \ref{eq:J} refers to the individual scalar components of the residual, solution, and Jacobian.
That is, for 3-D analyses, the residual takes the form
\begin{eqnarray}
	\bb{R}(\bb{u}) = \left\{ R_1^x, R_1^y, R_1^z, R_2^x, R_2^y, R_2^z, ..., R_n^z \right\}
\end{eqnarray}
and the solution takes the form
\begin{eqnarray}
	\bb{u} = \left\{ u_1^x, u_1^y, u_1^z, u_2^x, u_2^y, u_2^z, ..., u_n^z \right\}
\end{eqnarray}
In practice, it is often more practical and efficient to form and store the residual, solution and Jacobian in a \emph{blocked} manner, where the residual and solution can be considered as vectors of vectors.
Similarly, the Jacobian can be formed in terms of sub-matrix block coefficients.

@@

@@

To achieve this, the cell-centred displacement gradients are determined using a weighted first-neighbours least squares method \citep{Jasak1996},

@@

%		&=&\sum_{f_i \in \mathcal{F}^{\text{int}}_P} \frac{\bb{G}^{-1}_P \cdot \bb{d}_{f_i}}{\bb{d}_{f_i} \cdot \bb{d}_{f_i}}  \otimes \left(\bb{u}_{N_{f_i}} - \bb{u}_P \right) \notag \\
%		&&+ \sum_{d_i \in \mathcal{F}^{\text{disp}}_P} \frac{ \bb{G}^{-1}_P \cdot \bb{d}_{d_i} }{\bb{d}_{d_i} \cdot \bb{d}_{d_i}} \otimes \left(\bb{u}_{d_i} - \bb{u}_P \right) \notag \\
%		&&+ \sum_{s_i \in \mathcal{F}^{\text{symm}}_P} \frac{\bb{G}^{-1}_P \cdot \bb{d}_{s_i}}{\bb{d}_{s_i} \cdot \bb{d}_{s_i}} \otimes \left(\bb{R}_{s_i} \cdot \bb{u}_P - \bb{u}_P \right)
		&=&\sum_{f_i \in \mathcal{F}^{\text{int}}_P} w_{f_i} |\bb{\Gamma}_{f_i}| \frac{\bb{G}^{-1}_P \cdot \bb{d}_{f_i}}{\bb{d}_{f_i} \cdot \bb{d}_{f_i}}  \otimes \left(\bb{u}_{N_{f_i}} - \bb{u}_P \right) \notag \\
		&&+ \sum_{d_i \in \mathcal{F}^{\text{disp}}_P} w_{f_i} |\bb{\Gamma}_{d_i}| \frac{ \bb{G}^{-1}_P \cdot \bb{d}_{d_i} }{\bb{d}_{d_i} \cdot \bb{d}_{d_i}} \otimes \left(\bb{u}_{d_i} - \bb{u}_P \right) \notag \\
		&&+ \sum_{s_i \in \mathcal{F}^{\text{symm}}_P} w_{f_i} |\bb{\Gamma}_{s_i}| \frac{\bb{G}^{-1}_P \cdot \bb{d}_{s_i}}{\bb{d}_{s_i} \cdot \bb{d}_{s_i}} \otimes \left(\bb{R}_{s_i} \cdot \bb{u}_P - \bb{u}_P \right)

@@

%	 \sum_{{f_i} \in \mathcal{F}^{\text{int}}_P} \frac{\bb{d}_{f_i} \otimes \bb{d}_{f_i}}{\bb{d}_{f_i} \cdot \bb{d}_{f_i}}
%	 +  \sum_{{d_i} \in \mathcal{F}^{\text{disp}}_P} \frac{\bb{d}_{d_i} \otimes \bb{d}_{d_i}}{\bb{d}_{d_i} \cdot \bb{d}_{d_i}}
%	 +  \sum_{{s_i} \in \mathcal{F}^{\text{symm}}_P} \frac{\bb{d}_{s_i} \otimes \bb{d}_{s_i}}{\bb{d}_{s_i} \cdot \bb{d}_{s_i}}
	 \sum_{{f_i} \in \mathcal{F}^{\text{int}}_P} (1 - w_{f_i}) |\bb{\Gamma}_{f_i}|  \frac{\bb{d}_{f_i} \otimes \bb{d}_{f_i}}{\bb{d}_{f_i} \cdot \bb{d}_{f_i}} \notag \\
	 &&+  \sum_{{d_i} \in \mathcal{F}^{\text{disp}}_P} (1 - w_{d_i}) |\bb{\Gamma}_{d_i}|  \frac{\bb{d}_{d_i} \otimes \bb{d}_{d_i}}{\bb{d}_{d_i} \cdot \bb{d}_{d_i}} \notag \\
	 && +  \sum_{{s_i} \in \mathcal{F}^{\text{symm}}_P} (1 - w_{s_i}) |\bb{\Gamma}_{s_i}|  \frac{\bb{d}_{s_i} \otimes \bb{d}_{s_i}}{\bb{d}_{s_i} \cdot \bb{d}_{s_i}}

@@

As the $w |\bb{\Gamma}| \ \bb{G}^{-1}_P \cdot \bb{d}/(\bb{d}\cdot \bb{d})$ vectors in Equation \ref{eq:leastSquaresGrad} are purely a function of the mesh, they can be computed once and stored.

@@

first-order accurate gradients are sufficient to preserve second-order accuracy of the cell-centre displacements.

@@

So, this Rhie-Chow stabilisation can take the form of a \emph{jump} term; that is, it is a function of the jump in solution between neighbouring cells.
The two states, $\bb{u}_P^*$ and $\bb{u}_{N_{f_i}}^*$, are evaluated halfway between $P$ and $N_{f_i}$, and it may not be at a face centre; however, as noted by \citet{Nishikawa2010}, this will not affect the second-order accuracy of the method; consequently, the user is free to choose the value of $\alpha$ for stabilisation and accuracy purposes.
% but it doesn't matter because this team is only stabilisation and it vanishes in the grid refinement.
%As long as it vanishes for linear functions, second-order accuracy is preserved. It means that the parameter alpha can be chosen solely for the purpose of stabilization.

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Segregated Solution Algorithm} 
\label{sec:seg_alg}
%%--------------------------------------------------------------------------------------------------------------------%%
The classic segregated solution algorithm can be viewed as a quasi-Newton method, where an approximate Jacobian is derived from the inertia term and a compact-stencil discretisation of a diffusion term (from the stabilisation term):
%The surface forces Laplacian term (first term on the right-hand side of Equation \ref{eqn:MomentumImplicitExplicit}) is discretised using central differencing with over-relaxed non-orthogonal correction \cite{demirdzic_finite_1993, jasak_application_2000, cardiff_development_2014, cardiff_large_2014, cardiff_lagrangian_2017}:
\begin{eqnarray} \label{eq:diffusion}
	\tilde{\bb{J}} &=& \frac{\partial}{\partial \bb{u}} \left[ \oint_{\Gamma_P} \alpha \bar{K} \, \bb{n} \cdot \bb{\nabla} \bb{u} \; d\Gamma_P
	 \; -\;  \int_{\Omega_P} \rho \frac{\partial^2 \bb{u} }{\partial t^2} \, d\Omega_P \right]
\end{eqnarray}
The inertia term is discretised as described in Equations \ref{eq:inertia} and \ref{eq:inertia2}, while the diffusion term is discretised in the same manner as the compact-stencil component of the Rhie-Chow stabilisation term:
\begin{eqnarray}
	\oint_{\Gamma_P} \alpha \bar{K} \, \bb{n} \cdot \bb{\nabla} \bb{u} \; d\Gamma_P &\approx&
		\sum_{f_i \in \mathcal{F}^{\text{int}}_P}  \alpha \bar{K}
		\left|\bb{\Delta}_{f_i} \right| \frac{ \bb{u}_{N_{f_i}} - \bb{u}_P}{\left|\bb{d}_{f_i}\right|}    \left|\bb{\Gamma}_{f_i}\right| \notag \\
	&&+  \sum_{d_i \in \mathcal{F}^{\text{disp}}_P}  \alpha \bar{K}
%		\left|\bb{\Delta}_{d_i} \right| \left( \ \frac{ \bar{\bb{u}}_{d_i} - \bb{u}_P}{\left|\bb{d}_{d_i}\right|}  \right) 
		\left|\bb{\Delta}_{d_i} \right| \frac{ \bar{\bb{u}}_{d_i}  - \bb{u}_P}{\left|\bb{d}_{d_i}\right|} 
		\left|\bb{\Gamma}_{d_i}\right| \notag \\
	&&+ \sum_{s_i \in \mathcal{F}^{\text{symm}}_P}  \alpha \bar{K}
		\left|\bb{\Delta}_{s_i} \right| \frac{ \bb{R}_{s_i} \cdot \bb{u}_{P} - \bb{u}_P}{\left|\bb{d}_{s_i}\right|}
		\left|\bb{\Gamma}_{s_i}\right|
\end{eqnarray}
When a diffusion term is typically discretised using the cell-centre finite volume method, non-orthogonal corrections are included in a deferred correction manner to preserve the order of accuracy on distorted grids.
However, in the current quasi-Newton method form, the approximate Jacobian's exact value does not affect the final converged solution, but only the convergence behaviour.
Consequently, non-orthogonal corrections are not included in the approximate Jacobian here. However, grid distortion is appropriately accounted for in the calculation of the residual.
Nonetheless, as a result, it is expected that the convergence behaviour of the segregated approach may degrade as mesh non-orthogonality increases.
%Similarly, the prescribed displacements from the $\mathcal{F}^{\text{disp}}_P$ boundary faces does not affect the Jacobian.

The linearised system (Equation \ref{eq:Seg}) is formed for each cell in the domain, resulting in a system of algebraic equations:
\begin{eqnarray} \label{eq:SegSys}
    \left[ \bb{\tilde{J}} \right]  \; \left[ \delta \bb{u} \right] = - \left[\bb{R}(\bb{u}_k)\right]
\end{eqnarray}
where $\left[ \bb{\tilde{J}} \right]$ is a symmetric, $M \times M$ stiffness matrix, where $M = 3|\mathcal{P}|$ in 3-D and $M = 2|\mathcal{P}|$ in 2-D.
If $\Delta t < \infty$ or $\mathcal{F}^{\text{disp}}_P \neq \emptyset$, matrix $\left[ \bb{\tilde{J}} \right]$ is strongly diagonally dominant; otherwise, it is weakly diagonally dominant.
The block ($3\times3$ for 3-D, $2\times2$ for 2-D) diagonal coefficient for cell $P$ (row $P$, column $P$) can be expressed as
\begin{eqnarray}
	 \left[ \tilde{\bb{J}}\right]_{PP} &=&
		- \sum_{f_i \in \mathcal{F}^{\text{int}}_P}  \alpha \bar{K}
		\frac{ \left|\bb{\Delta}_{f_i} \right| }{\left|\bb{d}_{f_i}\right|}    \left|\bb{\Gamma}_{f_i}\right| \textbf{I} 
	    \quad-  \sum_{d_i \in \mathcal{F}^{\text{disp}}_P}  \alpha \bar{K}
%		\left|\bb{\Delta}_{d_i} \right| \left( \ \frac{ \bar{\bb{u}}_{d_i} - \bb{u}_P}{\left|\bb{d}_{d_i}\right|}  \right) 
		 \frac{ \left|\bb{\Delta}_{d_i} \right| }{\left|\bb{d}_{d_i}\right|} 
		\left|\bb{\Gamma}_{d_i}\right| \textbf{I} \notag \\
	 &&\quad - \sum_{s_i \in \mathcal{F}^{\text{symm}}_P} \alpha \bar{K}
		 \frac{ \left|\bb{\Delta}_{s_i} \right|}{\left|\bb{d}_{s_i}\right|}
		\left|\bb{\Gamma}_{s_i}\right|  \bb{n}_{s_i} \otimes \bb{n}_{s_i} 
	\quad - \quad \frac{9}{4}  \frac{\rho_P \Omega_P}{\Delta t^2} \textbf{I}
\end{eqnarray}
while the off-diagonal coefficients (row $P$, column $Q$) can be expressed as
\begin{eqnarray}
	\left[\tilde{\bb{J}}\right] _{PQ} &=&
		\alpha \bar{K} \frac{ \left|\bb{\Delta}_{f_{PQ}} \right| }{\left|\bb{d}_{f_{PQ}}\right|} \left|\bb{\Gamma}_{f_{PQ}}\right| \textbf{I} 
\end{eqnarray}
where subscript $PQ$ indicates a quantity associated with the internal face $f$ shared between cells $P$ and $Q$.
%where $N_f$ represents the set of faces $f$ in cell $P$, and neighbouring cell centre $N_f$ shares face $f$ with the cell $P$.
%The over-relaxed orthogonal vector $\bb{\Delta}_f = \frac{\bb{d}_f}{\bb{d}_f \cdot \bb{n}_f}$ 
%The non-orthogonal correction vector $\bb{k}_f=\bb{n}_f-\bb{\Delta}_f$.
% where $\bb{n}_f$ is the outward-facing unit normal to the face $f$.
%Vector $\bb{d}_f$ connects the centre of cell $P$ with the centre of cell $N_f$ in the updated configuration.
%The first term on the right-hand side is treated implicitly, while the second term - representing non-orthogonal corrections at the face - is treated in a deferred correction manner.

@@

In literature, the segregated solution algorithm is typically formulated in terms of the total displacement vector (or its difference between time steps) as the primary unknown; in contrast, in the quasi-Newton interpretation presented here, the primary unknown is the correction to the displacement vector, which goes to zero at convergence.
In the total displacement approach, the matrix is the same, and the only difference is the formulation of the right-hand side, which is calculated as the difference between the residual and the matrix times the previous solution.
Equivalence of both approaches can be seen by adding $\left[ \bb{\tilde{J}} \right]  \; \left[ \bb{u} \right]_k$ to both sides of Equation \ref{eq:SegSys}:
\begin{eqnarray} \label{eq:SegSys}
    \left[ \bb{\tilde{J}} \right]  \; \left[ \delta \bb{u} \right] + \left[ \bb{\tilde{J}} \right]  \; \left[ \bb{u} \right]_k  = - \left[\bb{R}(\bb{u}_k)\right] + \left[ \bb{\tilde{J}} \right]  \; \left[ \bb{u} \right]_k \notag \\
    \left[ \bb{\tilde{J}} \right]  \; \left[ \bb{u} \right]_{k+1} = \left[ \bb{\tilde{J}} \right]  \; \left[ \bb{u} \right]_k - \left[\bb{R}(\bb{u}_k)\right] \
\end{eqnarray}
where $\left[ \bb{u} \right]_k$ is the solution field at the previous iteration, and noting that $\left[ \bb{u} \right]_{k+1} = \left[ \bb{u} \right]_k + \left[ \delta \bb{u} \right]$.
In practice, the $\left[ \bb{\tilde{J}} \right]  \; \left[ \bb{u} \right]_k$ term on the right-hand side can be evaluated directly in terms of the discretised temporal and diffusion terms, without the need for matrix multiplication.
In the current work, the standard total displacement segregated formulation is employed.
%It is not expected that either formulation displays performance advantages.

As a final comment, the $\alpha$ scaling factor in the stabilisation term (Equation \ref{eq:RhieChow}) need not take the same value as in the approximate Jacobian (Equation \ref{eq:diffusion}).
In the current work, $\alpha$ is taken as unity in the approximate Jacobian, while the variation of its value in the residual calculation is examined in Section \ref{sec:RhieChowResults}.

@@

%\hl{3}
%- we use approx J to form precon
%- many precon used in lit, we will consider ILU($k$) and MG, where MG is expected to be better, but also LU since direct solvers are popular in FE solid mechanics
In the current work, we propose to use the compact-stencil approximate Jacobian from the segregated algorithm $\tilde{\bb{J}}$ as the preconditioning matrix $\bb{P}$ for the preconditioned Jacobian-free Newton-Krylov method.
%\hl{Comment on literature: used before but not yet for solids}.
This preconditioning approach can be considered as a ``physics-based" preconditioner in the classifications of \citet{Knoll2004}.
The approach is conceptually similar to an approximation of the Jacobian of a higher-order large-stencil scheme by a compact-stencil lower-order scheme.
A benefit of the proposed approach is that existing segregated frameworks can re-use their existing matrix assembly and storage implementations.
Concretely, the Jacobian-free Newton-Krylov method requires only a procedure for forming this preconditioning matrix and a procedure for explicitly evaluating the residual.
Both routines are easily implemented -- and are likely already available -- in an existing segregated framework.
The only additional required procedure is an interface to an existing Jacobian-free Newton-Krylov implementation.
%"The motivation behind this approach is that there exist numerous, legacy algorithms to solve nonlinear systems, both IVPs and BVPs. These algorithms typically were developed with some insight into the time scales or physical behavior of the problem. As a benefit of this insight, a reduced implicit system, or a sequence of segregated explicit or implicit systems may be solved in place of the fully coupled system. "
In the current work, the PETSc package (version 3.22.2) \cite{PETSc} is used as the nonlinear solver, driven by a finite volume solver implemented in the solids4foam toolbox \citep{Cardiff2018, Tukovic2018} for OpenFOAM toolbox \citep{Weller1998} (version OpenFOAM-v2312).

@@

%\hl{5 - oversolving}
%Comment on over-solving => also applies to the segregated system.
%This could be a parameter we look at.
A final comment on the Jacobian-free Newton-Krylov solution algorithm is the potential importance of \emph{oversolving}.
Here, oversolving refers to solving the linear system to a tolerance that is too tight during the early Newton iterations, essentially wasting time when the solution is far from the true solution.
In addition, some authors %\cite{See164_and_176_in_Knoll2004}
\cite{Knoll2004} have shown Newton convergence to be worse when the earlier iterations are solved to too tight a tolerance.
The concept of oversolving also applies to segregated solution procedures and has been well-known since the early work of Demird\v{z}i\'{c} and co-workers \cite{Demirdzic1995}, where the residuals are typically reduced by one order of magnitude in the inner linear system.
%The optimal choice of residual reduction for a Jacobian-free Newton-Krylov finite volume solid mechanics procedure is explored in Section \ref{sec:test_cases}. \hl{Check: do we examine this?}
In the current work, the residual is reduced by one order of magnitude in the segregated approach and three orders in the Jacobian-free Newton-Krylov approach.
%\hl{Try for EW plastic: $-snes_ksp_ew -snes_ksp_ew_version 2 -snes_ksp_ew_rtol0 1e-4$}

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Checking Convergence}
%%--------------------------------------------------------------------------------------------------------------------%%
The current work adopts three checks for determining whether convergence has been achieved within each time (or loading) step, closely following the default strategy in the PETSc toolbox:
\begin{itemize}
    \item Norm of the residual $|\bb{R}(\bb{u})|$:
    \begin{itemize}
        \item \emph{Absolute}: Convergence is declared when $|\bb{R}(\bb{u})|$ falls below an absolute tolerance $a_{\text{tol}}$, taken here as $1\times 10^{-50}$. This acts as a failsafe in situations where the \emph{relative} criterion (below) is ineffective because the initial solution already nearly satisfies the governing equation.
        \item \emph{Relative}: Convergence is declared when $|\bb{R}(\bb{u})|$ falls below $r_{\text{tol}} \times |\bb{R}(\bb{u}_0)|$, with $|\bb{R}(\bb{u}_0)|$ denoting the residual norm at the start of the time (or loading) step. In this work, $r_{\text{tol}} = 10^{-6}$.
    \end{itemize}
    \item Norm of the solution correction $|\delta \bb{u}|$: Convergence is declared when the norm of the change in the solution between successive iterations is less than a prescribed tolerance, $s_{\text{tol}}$.
\end{itemize}

Although it may seem reasonable to apply the same tolerances for both the segregated and Jacobian-free Newton-Krylov solution procedures, care must be taken with $s_{\text{tol}}$. The Jacobian-free Newton-Krylov method typically makes a small number of large corrections to the solution, whereas the segregated procedure often makes a large number of small corrections. Consequently, if the same $s_{\text{tol}}$ value were used, the segregated procedure might be declared convergent prematurely. To avoid this, $s_{\text{tol}}$ is set to zero in the present work, meaning convergence is based solely on residual reduction. Nonetheless, setting $s_{\text{tol}} > 0$ could yield a more robust procedure capable of handling potential stalling.
